Wiggling the bandwidth in ridgeless regression with scale-dependent kernels

### 2023-01-16 paper.tex
add bibliography and write interpolating regime motivations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let $\mathbf X \in \R^{n \times d}$ be a matrix of $n$ training examples
$X_1, \dots, X_n \in \R^d$ in $d$-dimensional Euclidean space as rows and 
let $Y$ be the column vector of responses $y_1, \dots, y_n \in \R$.
The main actor of this paper is the kernel ridgeless regression estimator
that predicts the response on an unseen data point $X \in \R^d$ as
\begin{equation} \label{eq:ridgeless}
    \hat f(X) = \sum_{i=1}^n (\mathbf K^\dagger Y)_i K(X, X_i).
\end{equation}
Here, $mathbf K$ denotes the kernel matrix $\mathbf K_{ij} = k(X_i, X_j)$
is the \emph{kernel matrix} obtained by evaluating a kernel function 
$k \colon \R^d \times \R^d \to \R$ on the training examples and 
$\mathbf K^\dagger$ denotes its \emph{pseudo-inverse}.
We briefly recall the motivations behind these notions.
First, the \emph{representer theorem} asserts that any minimizer to the 
least squares problem
\[ \min_{f \in \mathcal H_k} \frac 1n \sum_{i=1}^n (y_i - f(X_i))^2 \]
in the \emph{reproducing kernel Hilbert space} $\mathcal H_k$ associated
to $k$, is given by a linear combination of the form
$\sum_{i=1}^n c_i k(X,X_i)$.
This allows to apply the usual linear least squares method to a - at the 
first glance - non-linear approximation problem in a possibly 
infinite-dimensional hypothesis space by pushing the data points into 
the \emph{feature space} spanned by $K(X, \cdot) \in \mathcal H_k$ 
for $X \in \R^d$.
The kernel matrix in this setting corresponds to the empirical covariance
matrix $\mathbf X^T \mathbf X$ and classical theory suggests that, even if 
it does not possess full-rank, there will always exist a minimizer to 
the least squares problem above.
The pseudo-inverse then corresponds to the unique least-squares solution
with minimal $\mathcal H_k$-norm.
This reduction of an infinite-dimensional optimization problem to a
finite-dimensional one using kernels is generally known as the 
\emph{kernel trick}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Wiggling as a regularization technique}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Iterative wiggling to find the best bandwidth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and directions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Let $\mathbf X \in \R^{n \times d}$ be a matrix of $n$ training examples
$X_1, \dots, X_n \in \R^d$ in $d$-dimensional Euclidean space as rows and 
let $Y$ be the column vector of responses $y_1, \dots, y_n \in \R$.
The main actor of this paper is the kernel ridgeless regression estimator
that predicts the response on an unseen data point $X \in \R^d$ as
\begin{equation} \label{eq:ridgeless}
    \hat f(X) = \sum_{i=1}^n (\mathbf K^\dagger Y)_i K(X, X_i).
\end{equation}
Here, $\mathbf K$ denotes the kernel matrix $\mathbf K_{ij} = k(X_i, X_j)$
is the \emph{kernel matrix} obtained by evaluating a kernel function 
$k \colon \R^d \times \R^d \to \R$ on the training examples and 
$\mathbf K^\dagger$ denotes its \emph{pseudo-inverse}.
We briefly recall the motivations behind these notions.
First, the \emph{representer theorem} asserts that any minimizer to the 
least squares problem
\[ \min_{f \in \mathcal H_k} \frac 1n \sum_{i=1}^n (y_i - f(X_i))^2 \]
in the \emph{reproducing kernel Hilbert space} $\mathcal H_k$ associated
to $k$, is given by a linear combination of the form
$\sum_{i=1}^n c_i k(X,X_i)$.
This allows to apply the usual linear least squares method to a - at the 
first glance - non-linear approximation problem in a possibly 
infinite-dimensional hypothesis space by pushing the data points into 
the \emph{feature space} spanned by $K(X, \cdot) \in \mathcal H_k$ 
for $X \in \R^d$.
The kernel matrix in this setting corresponds to the empirical covariance
matrix $\mathbf X^T \mathbf X$ and classical theory suggests that, even if 
it does not possess full-rank, there will always exist a minimizer to 
the least squares problem above.
The pseudo-inverse then corresponds to the unique least-squares solution
with minimal $\mathcal H_k$-norm.
This reduction of an infinite-dimensional optimization problem to a
finite-dimensional one using kernels is generally known as the 
\emph{kernel trick}.

Recently, there was a new wave of interest into kernels sparked by the 
fundamental question of why highly over-parameterized deep neural nets
perform as well as they do given that the usual established model, 
the \emph{bias-variance trade-off}, does not capture this regime.
The idea is that to find a good estimator one should choose the 
complexity of the hypothesis space should be high enough to avoid 
\emph{underfitting} but simultaneously low enough to avoid 
\emph{overfitting}.
For instance, if we roughly identify the number of adjustable parameters
of a model with the complexity of the hypothesis space, we would expect
a U-shaped test error curve over the number of parameters with very high
risk at the extrema where $N_{\textrm{param}}$ is very small but also
$N_{\textrm{param}} \approx N_{\textrm{train}}$, where the model is able
to \emph{interpolate} the training data - it learns them by heart and 
the analogy is the same as in the human experience where learning by heart
does not give you true understanding.
However, modern, practically successful, deep neural nets tend to have a
parameter count that exceeds the training sample count by orders of 
magnitude.
This is not captured by classical theory so the question is two-fold:
is this the reason for their performance and 
what exactly happens in this \emph{interpolating regime}.
A good start into the literature would be the following papers on 
the \emph{double descent} curve (\cite{doubledescent}) and on the new 
concept of \emph{benign overfitting} (\cite{benignoverfitting}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Wiggling as a regularization technique}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Iterative wiggling to find the best bandwidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and directions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{20}

\bibitem{doubledescent}
Belkin, Mikhail, et al. "Reconciling modern machine learning practice and the bias-variance trade-off." arXiv preprint arXiv:1812.11118 (2018).

\bibitem{benignoverfitting}
Bartlett, Peter L., et al. "Benign overfitting in linear regression." Proceedings of the National Academy of Sciences 117.48 (2020): 30063-30070.

\bibitem{understandkernels}
Belkin, Mikhail, Siyuan Ma, and Soumik Mandal. "To understand deep learning we need to understand kernel learning." International Conference on Machine Learning. PMLR, 2018.

\bibitem{justinterpolate}
Liang, Tengyuan, and Alexander Rakhlin. "Just interpolate: Kernel “ridgeless” regression can generalize." The Annals of Statistics 48.3 (2020): 1329-1347.

\end{thebibliography}
\end{document}
